InformatiCup 2024 Ctrl-Alt-Defeat

### 2024-01-14 paper.tex
Apply suggestions from Lukas
Co-authored-by: EhrenbruderRoschtbraten <81183139+EhrenbruderRoschtbraten@users.noreply.github.com>

The first image generative model we use is the Stable Diffusion model. It is developed by engineers and researchers from CompVis (\url{https://github.com/CompVis}), Stability AI (\url{https://stability.ai/}) and LAION (\url{https://laion.ai/}) \cite{patil2022stable}.

Diffusion models achieve state-of-the-art synthesis results on image data by decomposing the image formation process into sequential applications of denoising autoencoders. Due to its work directly in pixel space, they are very computational expensive. To lower the computational costs, the Stable Diffusion model (v1.5) is applied in a latent space of powerful pretrained autoencoders. This works without losing quality and flexibility. \cite{Rombach_2022_CVPR}

As a second image generative model we use OpenDalle from dataautogpt3 (\url{https://huggingface.co/dataautogpt3/OpenDalle}). OpenDalle, as it is based on the Dall-E model, works by modelling text and image tokens as a single data stream. This is achieved by training a transformer to autoregressively model those tokens into the resulting stream.

This is done in two steps: 
\begin{enumerate}
	\item The images get compressed in a $32 \times 32$ grid of image tokens to reduce the context size of the transformer.
	\item BPE-encoded text tokens are concatenated with the image tokens to model a joint distribution over text and image tokens with the training of an autoregressive transformer.
\end{enumerate}

The exact procedure how DALL-E models work proposed by Ramesh et. al is explained in their paper. \cite{ramesh2021zeroshot}

@@

\section{Results}

\section{Discussion}

>>>>

@@

\section{Results + Discussion}
In the following section we take a look on the results of the aforementioned experiment.

Fig. \ref{stable_diffusion} shows how the different processors and their combinations effect an image generated by the stable diffusion generator. The bars are structured in packs of three where the most left one represents the nahrawy evaluator, the middle one is resnet18 and on the right is the umm-maybe evaluator respectively. The different colors represent the different outputs. Black symbolizes that the image before and after processing was detected as AI-generated. Grey shows that both are detected to be human made, whereas for red the process did the exact opposite of what it should do and fool the detector in a way that after processing it is seen as AI-generated whereas before as human made. Lastly the green bar does exactly what is desired, so it pictures AI-generated content, where the executed augmentation lead to the detector guessing it was human-made. \\
The data depicted in fig. \ref{fig:stable_diffusion} shows, that only the nahrawy evaluator fooled the detector the wrong way round, but it is also the only one with the output of 'human-human'. Compared to the Dall-E generator in fig. \ref{fig:dallE} the amount of right tricking is relatively low. There are also other differences as here only umm-maybe achieved 'human-ai' but also 'human-human'. 
It is visible that the gaussian noise image processor as well as the salt-and-pepper one are the best processors. This can be seen as they solely achieve a great result as well as their combinations with other ones. It is also worth mentioning that these results are only achieved when using the nahrawy evaluator. Whereas resnet18, which can be seen in both figures, couldn't be tricked by any processor.

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{images/stable-diffusion.png}
	\caption{Stable Diffusion with different processors and nahrawy, resnet18 and umm-maybe as evaluators}
	\label{fig:stable_diffusion}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{images/dalle.png}
	\caption{Dall-E with different processors and nahrawy, resnet18 and umm-maybe as evaluators}
	\label{fig:dallE}
\end{figure}

In fig. \ref{fig:gpt2} the text processors are plotted. The colors stay the same and the concept as well, where the left bar is using the radar evaluator and the right one the roberta base one. It is visible that the double-whitespace processor solely has little effect whereas in combination with the typo processor the result rises enormously. This is due to the typo processor since it alone also achieves great results. The radar evaluator only produces only black bars. 
Similar results can be seen in fig. \ref{fig:falcon} where radar just contains black bars and typo as well, as its combination with the roberta base evaluator achieves great results. The difference is that with the falcon generator and the roberta base evaluator in combination, some texts are detected as human made even before processing, which explains the red and grey parts.

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{images/gpt2.png}
	\caption{GPT2 with different processors and nahrawy, resnet18 and umm-maybe as evaluators}
	\label{fig:gpt2}
\end{figure}

\begin{figure}
	\centering
	\includegraphics[width=\linewidth]{images/falcon.png}
	\caption{Falcon with different processors and nahrawy, resnet18 and umm-maybe as evaluators}
	\label{fig:falcon}
\end{figure}

To summarize we can say, that the combination between generators and evaluators plays a huge role in the results. In case of the text generators the difference seems minor, but for example the roberta base evaluator detected around ten falcon generated texts as human made in the middle case of fig. \ref{fig:falcon} and after processing as well, which means that if that would not have been the case, we would have a perfect score here. 
This becomes more clear when comparing fig. \ref{fig:stable_diffusion} with fig \ref{fig:dallE} where the importance of the used generators becomes visible. The stable-diffusion generator produces images where it is difficult to fool a detector whereas when using the same processors and detectors, the images of dallE can be processed in a way to fool some detectors.

Another noticeable result is in the usage of different evaluators. In both cases, text and image, we had one evaluator respectively which could not be fooled with any method. These evaluators where so strict, that they would even classify a human made image provided by us as AI-generated.
